Merging Datasets Through Deep learning

### 2018-08-29 paper2018/introduction2.tex
adding more text

To test the feasability of these ideas, we used Wikidata as ground truth to build models for datasets with 1.1M people's names (about 200K identities) and 130K company names (70K identities).  The problem of mapping vectors for the same entity closer in vector space than vectors for other entities is known in the literature as deep metric learning.  Deep metric learning is known to be a difficult problem as studied in the space of face recognition and person-re-identification \cite{DBLP:conf/cvpr/SchroffKP15}.  As a result, there is a significant amount of research that has been targeted at two aspects of training these networks (a) how to choose samples for efficient learning, and (b) what constitutes a good loss function.  In building models for entity names, we developed a novel method for sample selection, and a novel loss function, which was effective in building these models.

As in face recognition, our system for metric learning is built by training a so-called `siamese triplet' network to learn to produce a small distance estimate for two surface forms for the same entity (positive pairs), and a large distance estimate for surface forms of different entities (negative pairs).  A key problem in effective training of such networks is the problem of how to select negative pair for training, because one cannot exhaustively show all negative pairs to the network.  In prior work for instance, this problem is solved by so-called `triplet mining' where negatives are gleaned after an all-pairs comparison of input vectors in a batch \cite{DBLP:conf/cvpr/SchroffKP15}.  Our novel approach to this problem is to re-use the approximate nearest neighbors algorithm to choose negatives for training.  Because our ultimate goal is to train the system to learn to map positives pairs closer and negative pairs further away from each other, we select negatives that appear in the nearest neighbor set for input vectors of a given entity, before any training has taken place.  This approach has two key benefits.  First, it lays out the entities based on their input vectors, and thus allows an efficient gathering of all nearest neighbors without a quadratic comparison of vectors to determine suitable negatives.  Second, it provides a baseline against which one can objectively measure the effects if any of training.  If for instance, the input vectors for surface forms of the same entity are already close together in vector space, little can be gained by a deep learning model.

We also investigated the effect of manipulating multiple loss functions which have been proposed in the literature for improving deep metric learning, e.g., the triplet loss \cite{DBLP:conf/cvpr/SchroffKP15}, improved loss \cite{Zhang:2016:DML:3088616.3088665}, and angular loss \cite{DBLP:journals/corr/abs-1708-01682} functions.  For the problem of building models for people's names, we found that a modification to the objective function proposed in \cite{DBLP:conf/cvpr/SchroffKP15} was better than three other functions that have been proposed in the literature.

>>>>

To test the feasability of these ideas, we used Wikidata as ground truth to build models for datasets with 1.1M people's names (about 200K identities) and 130K company names (70K identities).  The problem of mapping vectors for the same entity closer in vector space than vectors for other entities is known in the literature as deep metric learning.  Deep metric learning is known to be a difficult problem as studied in the space of face recognition and person-re-identification \cite{DBLP:conf/cvpr/SchroffKP15}.  As a result, there is a significant amount of research that has been targeted at two aspects of training these networks (a) how to choose samples for efficient learning, and (b) what constitutes a good loss function.  In building models for entity names, we had to adapt these techniques for triplet selection and loss functions because matching entity names have different properties than faces, as we describe below.

As in face recognition, our system for metric learning is built by training a so-called triplet based `siamese triplet' network to learn to produce a small distance estimate for two surface forms for the same entity (between an arbitarily chosen anchor and positive), and a large distance estimate for surface forms of different entities (an anchor and a negative).  A key problem in effective training of such networks is the problem of how to select negative pair for training, because one cannot exhaustively show all negative pairs to the network.  In prior work for instance, this problem is solved by so-called `triplet mining' where so-called `semi-hard' negatives are gleaned after an all-pairs comparison of input vectors in a batch, e.g., \cite{DBLP:conf/cvpr/SchroffKP15}.  `Semi-hard' negatives are negatives that are further away from the anchor than positives, but not by a sufficient margin.  The idea of focusing on semi-hard negatives is to avoid mining poorly structured regions of the embedding space.  For matching across entity names, most positive forms for the entity names are actually substantially further away from the anchor and from negatives, as we show empirically in this paper.  Our approach to this problem is to use an approximate nearest neighbors algorithm to choose negatives for training in an efficient way, which means that most of the time, our negatives are `hard negatives', where the negative is closer to the anchor than positives.  This approach has three key benefits.  First, it lays out the entities based on their input vectors, and thus allows an efficient gathering of all nearest neighbors without a quadratic comparison of vectors in a batch to determine suitable negatives.  Second, it provides a baseline against which one can objectively measure the effects if any of training.  If for instance, the input vectors for surface forms of the same entity are already close together in vector space, little can be gained by a deep learning model.  Third, it examines whether focusing on hard negatives is really detrimental to deep metric learning, at least for learning to map entity names in vector space.

We also investigated the effect of manipulating multiple local loss functions which have been proposed in the literature for improving deep metric learning, e.g., the triplet loss \cite{DBLP:conf/cvpr/SchroffKP15}, improved loss \cite{Zhang:2016:DML:3088616.3088665}, and angular loss \cite{DBLP:journals/corr/abs-1708-01682} functions.  For the problem of building models for people's names, we found that a modification to the objective function proposed in \cite{DBLP:conf/cvpr/SchroffKP15} was better than three other functions that have been proposed in the literature.

### 2018-08-29 paper2018/main.tex
adding more text

\maketitle
\begin{abstract}
Merging datasets is a key operation for data analytics.  A frequent requirement for merging is joining across columns that have different surface forms for the same entity (e.g., the name of a person might be represented as \textit{Douglas Adams} or \textit{Douglas Noel Adams}).  Data management systems are currently limited to performing merges based on string equality, or at best use string similarity.  We propose a novel approach to performing merges based on deep learning models.  Our approach depends on (a) creating a deep learning model that maps surface forms of an entity into a set of vectors such that alternate forms for the same entity are closest in vector space, (b) indexing these vectors using a nearest neighbors algorithm to find the forms that can be potentially joined together.  Using Wikidata as our ground truth, we built models for datasets with approximately 1.1M people's names (200K identities) and 130K company names (70K identities), where the vector embeddings allow for joins with precision of .91-.93 and recall of .70-.81. To build these models, we developed novel techniques for the selection of samples to train on, as well as a loss function, which we describe in the paper.  We make the models available to the community for aligning people or companies across multiple datasets.

>>>>

\maketitle
\begin{abstract}
Merging datasets is a key operation for data analytics.  A frequent requirement for merging is joining across columns that have different surface forms for the same entity (e.g., the name of a person might be represented as \textit{Douglas Adams} or \textit{Douglas Noel Adams}).  Data management systems are currently limited to performing merges based on string equality, or at best use string similarity.  We propose a novel approach to performing merges based on deep learning models.  Our approach depends on (a) creating a deep learning model that maps surface forms of an entity into a set of vectors such that alternate forms for the same entity are closest in vector space, (b) indexing these vectors using a nearest neighbors algorithm to find the forms that can be potentially joined together.  Using Wikidata as our ground truth, we built models for datasets with approximately 1.1M people's names (200K identities) and 130K company names (70K identities), where the vector embeddings allow for joins with precision of .91-.93 and recall of .70-.81. To build these models, we adapted techniques for the selection of samples to train on, as well as a loss function, which we describe in the paper.  We make the models available to the community for aligning people or companies across multiple datasets.

### 2018-08-29 paper2018/relatedWork.tex
adding more text

TBD - related work on triplet selection and losses.

>>>>

The use of nearest neighbors algorithms for triplet mining has been applied by \cite{DBLP:journals/corr/KumarHC0D17}, but their focus is on mining semi-hard negatives, that is focusing on negatives where at least one positive is closer to the anchor.  In general, this means that one cannot look at fixed neighborhood sizes in building triplets.  In other words, if all the positives are further away from the anchor than the negatives in a given neighborhood size of $k$, it means that $k$ needs to be expanded until a neighborhood size is found that has the right characteristics (i.e., at least one positive that is closer than the negattives which can be chosen).   Our approach much more efficient because it relies on a fixed $k$ to generate samples.
TBD - related work on triplet selection and losses.
